AMMBR: A Method to Maintain Business Rules

### 2022-09-05 AMMBR/patent.tex
Include the AMMBR idea for future publication

\documentclass[11pt,a4paper,fleqn,oneside]{article}
\usepackage{amssymb}
\parskip 10pt plus 2.5pt minus 4pt  % Extra vertical space between paragraphs.
\parindent 0em                      % Width of paragraph indentation.
\addtolength{\textwidth}{.5in}
\newcommand{\tick}{\mbox{\(\surd\)}}
\long\def\@savemarbox#1#2{\global\setbox#1\vtop{\hsize\marginparwidth 
   \@parboxrestore\smal #2}}
\def\bottomfraction{.5}
\renewcommand{\textfraction}{0.2}
\renewcommand{\floatpagefraction}{0.7}
\setcounter{tocdepth}{1}
\makeatletter
\def\@dummydocumentstyle{\@ifnextchar[{\@@dds}{\@@dds[]}}
\def\@@dds[#1]#2{\relax}

\def\getchapter#1{{\def\documentstyle{\@dummydocumentstyle}%
		   \def\document{\relax}%
		   \def\enddocument{\relax}\include{#1}}} 
\makeatother
\newlength{\Example}\setlength{\Example}{\textwidth}
\addtolength{\Example}{-0.5in}
\newlength{\Exinside}\setlength{\Exinside}{\Example}
\addtolength{\Exinside}{-0.5in}
\def\theor{\list{}{\leftmargin .5in}\item[]{\ }\\*[-\itemsep] \rm}
\let\endtheor=\endlist
\newtheorem{lemma}[equation]{Lemma}
\newtheorem{define}[equation]{Definition}
\newtheorem{examp}[equation]{Example}
%% end taken from preamble
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\def\rbracket{\text{]}}
\def\lbracket{\text{[}}
%\newcommand{\powerset}[1]{{\cal P}\{#1\}}
\newcommand{\one}[1]{{\cal\bf 1}_{#1}}
\newcommand{\choice}[2]{#1\ |\ #2}
\newcommand{\all}[2]{#1 + #2}
\newcommand{\seq}[2]{#1 ; #2}
\newcommand{\eca}[2]{{\tt ON}~#1~{\tt DO}~#2}
\newcommand{\subst}[3]{#3_{\tt [#1 := #2]}}
\newcommand{\assign}[2]{#1~{\tt :=}~#2}
\newcommand{\code}[1]{\text{\tt\small #1}}
\newcommand{\relAdd}{\dagger}
\newcommand{\dom}[1]{\id{dom}_#1}
\newcommand{\cod}[1]{\id{cod}_#1}
\newcommand{\conts}[1]{\id{contents}_#1}
\newcommand{\name}[1]{\id{name}_#1}
\def\id#1{\text{\em #1\/}}
\def\define#1{\label{dfn:#1}{\em #1}\index{#1}}
\def\definem#1{\label{dfn:#1}{\em #1}\index{#1}\marginpar{\flushright\small\em #1}}
\newcommand{\marg}[1]{\index{#1}\marginpar{\flushright\small\em #1}}
\newcommand{\uni}{\text{\small UNI}}
\newcommand{\tot}{\text{\small TOT}}
\newcommand{\inj}{\text{\small INJ}}
\newcommand{\sur}{\text{\small SUR}}
\newcommand{\univalent}[1]{\id{univalent}(#1)}
\newcommand{\total}[1]{\id{total}(#1)}
\newcommand{\surjective}[1]{\id{surjective}(#1)}
\newcommand{\injective}[1]{\id{injective}(#1)}
\newcommand{\transitive}[1]{\id{transitive}(#1)}
\newcommand{\symmetric}[1]{\id{symmetric}(#1)}
\newcommand{\antisymmetric}[1]{\id{antisymmetric}(#1)}
\newcommand{\reflexive}[1]{\id{reflexive}(#1)}
\newcommand{\ident}[1]{\mathbb{I}_{#1}}
\newcommand{\iden}{\mathbb{I}}
\newcommand{\full}{\mathbb{V}}
\newcommand{\flip}[1]{{#1}^\smallsmile}
\newcommand{\kleeneplus}[1]{{#1}^+}
\newcommand{\kleenestar}[1]{{#1}^*}
\newcommand{\cmpl}[1]{\overline{#1}}
\newcommand{\rel}{\times}
\newcommand{\compose}{;}
\newcommand{\subs}{\subseteq}
\newcommand{\fun}{\rightarrow}
\newcommand{\N}{\mbox{\msb N}}
\newcommand{\disjn}[1]{\id{disjoint}(#1)}
\newcommand{\fsignat}[3]{\id{#1}:\id{#2}\mbox{$\rightarrow$}\id{#3}}
\newcommand{\signat}[3]{\mbox{${#1}_{[{#2},{#3}]}$}}
\newcommand{\declare}[3]{\id{#1}:\id{#2}\mbox{$\times$}\id{#3}}

\title{AMMBR: A Method to Maintain Business Rules}
\author{Stef Joosten}
\begin{document}
\maketitle
\begin{abstract} % 149 words counted.
%<claim>
	This paper proposes a method, AMMBR, to generate software for maintaining the transactional integrity of a dataset.
%</claim>
	The purpose of this method is to enhance development environments that generate information systems with the feature of built-in compliance.
	Inspiration for this problem came from the author's practice of designing business processes in organisations.
	The idea is to let a dataset represent the facts of an organization,
	and let each business rule correspond to one invariant on that dataset.
	Business rules can be maintained by using software to detect violations of the invariants.
	That software can be generated in full by a generator, which analyses the invariants and compiles them into computer code.
	Thus, AMMBR determines the computer code to restore the invariants, when broken by events that change the dataset.
	This method has been tried and implemented in Ampersand, a tool that generates functional specifications and information systems.
\end{abstract}
keywords: compliance, business rule, relation algebra, software generation, provable software, process control.
\section{Introduction}
	In order to define an information system to support an organization,
	we look at organizations as systems that maintain rules.
	An ideal information infrastructure supports employees and other stakeholders to maintain the rules of the business.
	Yet rules must be concrete if they are to be used by computers.
	The method is therefore restricted to those rules of which
	violation can be established unambiguously and objectively.
	To {\em maintain} a rule means to prevent or correct all violations, either by human or automated actions.
	Temporary violation of a rule is only acceptable for rules that are maintained by humans,
	but the infrastructure must signal these violations and make them available to those people

@@

\section{Clauses}
\label{sct:Clauses}
	This section explains:
\begin{enumerate}
\item	how to derive clauses from a set of rules, by a normalization procedure.
\item	how to transform clauses into left identity clauses and right identity clauses, if possible.
\item	how to derive multiplicity properties from clauses thus obtained.
\end{enumerate}
	A \define{rule} is an expression $e$ that must be kept true.
	To \define{maintain} rule $r$ means to ensure that $\full\subs r$ is true,
	where $\full$ is the complete relation (i.e. everything is an pair of $\full$).
	In order to discuss maintaining rules, we must introduce some vocabulary.
	As long as $\full\subs r$, we say that $r$ is \define{satisfied}, or simply that it \define{holds}.
	If $r$ is not $\full$, we say that \define{invariance is broken},
	or rule $r$ is \define{not satisfied}, or it \define{does not hold}.
	Since $r\cup\cmpl{r}=\full$, the missing pairs are determined by $\cmpl{r}$.
	Each pair of $\cmpl{r}$ is called a \define{violation} of $r$.
	To \define{restore invariance} of rule $r$ means to change the contents of $r$ in such a way that $\full\subs r$.

@@

\subsection{old stuff}

@@

	Each of the code fragments from table \ref{tbl:code fragments} will preserve invariance of $e$.
	We shall prove this for one code fragment only (we have arbitrarily picked the third code fragment):
	This gives
\[\begin{array}{rl}
&\cmpl{(p\cup\Delta)}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ (t\cup\cmpl{a(e)})\\
=&~~~~~~~~\{a(e)\ =\ \cmpl{(p\cup\Delta)}\cup\cmpl{q}\cup s\cup t\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{\cmpl{(p\cup\Delta)}\cup\cmpl{q}\cup s\cup t}\\
=&~~~~~~~~\{\text{De Morgan (\ref{eqn:De Morgan})}\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ (\cmpl{\cmpl{(p\cup\Delta)}}\cap\cmpl{\cmpl{q}}\cap\cmpl{s}\cap\cmpl{t})\\
=&~~~~~~~~\{\text{remove double complement}\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ ((p\cup\Delta)\cap q\cap\cmpl{s}\cap\cmpl{t})\\
=&~~~~~~~~\{\text{Distribute $\cap$ over $\cup$}\}\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ (p\cup\Delta))\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ q)\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{s})\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{t})\\
=&~~~~~~~~\{\cmpl{p\cup\Delta}\ \cup\ p\cup\Delta=\full\text{~and~}\cmpl{q}\cup q=\full\text{~and~}s\ \cup\ \cmpl{s}=\full\text{~and~}t\ \cup\ \cmpl{t}=\full\}\\
&\full\cap \full\cap \full\cap \full\\
=\\
&\full\\
\end{array}\]

>>>>

\documentclass[11pt,a4paper,fleqn,oneside]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\parskip 10pt plus 2.5pt minus 4pt  % Extra vertical space between paragraphs.
\parindent 0em                      % Width of paragraph indentation.
\addtolength{\textwidth}{.5in}
\newcommand{\tick}{\mbox{\(\surd\)}}
\long\def\@savemarbox#1#2{\global\setbox#1\vtop{\hsize\marginparwidth 
   \@parboxrestore\smal #2}}
\def\bottomfraction{.5}
\renewcommand{\textfraction}{0.2}
\renewcommand{\floatpagefraction}{0.7}
\setcounter{tocdepth}{1}
\makeatletter
\def\@dummydocumentstyle{\@ifnextchar[{\@@dds}{\@@dds[]}}
\def\@@dds[#1]#2{\relax}

\def\getchapter#1{{\def\documentstyle{\@dummydocumentstyle}%
		   \def\document{\relax}%
		   \def\enddocument{\relax}\include{#1}}} 
\makeatother
\newlength{\Example}\setlength{\Example}{\textwidth}
\addtolength{\Example}{-0.5in}
\newlength{\Exinside}\setlength{\Exinside}{\Example}
\addtolength{\Exinside}{-0.5in}
\def\theor{\list{}{\leftmargin .5in}\item[]{\ }\\*[-\itemsep] \rm}
\let\endtheor=\endlist
\newenvironment{derivation}{\begin{tabbing}\advance\mathindent -1em\hspace*{1em}\=\hspace*{\mathindent}\=\hspace*{.6in}\=etc.\+\kill}{\end{tabbing}}
\newtheorem{lemma}[equation]{Lemma}
\newtheorem{define}[equation]{Definition}
\newtheorem{examp}[equation]{Example}
%% end taken from preamble
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\def\rbracket{\text{]}}
\def\lbracket{\text{[}}
%\newcommand{\powerset}[1]{{\cal P}\{#1\}}
\newcommand{\one}[1]{{\cal\bf 1}_{#1}}
\newcommand{\choice}[2]{#1\ |\ #2}
\newcommand{\all}[2]{#1 + #2}
\newcommand{\seq}[2]{#1 ; #2}
\newcommand{\eca}[2]{{\tt ON}~#1~{\tt DO}~#2}
\newcommand{\subst}[3]{#3_{\tt [#1 := #2]}}
\newcommand{\assign}[2]{#1~{\tt :=}~#2}
\newcommand{\code}[1]{\text{\tt\small #1}}
\newcommand{\relAdd}{\dagger}
\newcommand{\dom}[1]{\id{dom}_#1}
\newcommand{\cod}[1]{\id{cod}_#1}
\newcommand{\conts}[1]{\id{contents}_#1}
\newcommand{\name}[1]{\id{name}_#1}
\def\id#1{\text{\em #1\/}}
\def\define#1{\label{dfn:#1}{\em #1}\index{#1}}
\def\definem#1{\label{dfn:#1}{\em #1}\index{#1}\marginpar{\flushright\small\em #1}}
\newcommand{\marg}[1]{\index{#1}\marginpar{\flushright\small\em #1}}
\newcommand{\uni}{\text{\small UNI}}
\newcommand{\tot}{\text{\small TOT}}
\newcommand{\inj}{\text{\small INJ}}
\newcommand{\sur}{\text{\small SUR}}
\newcommand{\univalent}[1]{\id{univalent}(#1)}
\newcommand{\total}[1]{\id{total}(#1)}
\newcommand{\surjective}[1]{\id{surjective}(#1)}
\newcommand{\injective}[1]{\id{injective}(#1)}
\newcommand{\transitive}[1]{\id{transitive}(#1)}
\newcommand{\symmetric}[1]{\id{symmetric}(#1)}
\newcommand{\antisymmetric}[1]{\id{antisymmetric}(#1)}
\newcommand{\reflexive}[1]{\id{reflexive}(#1)}
\newcommand{\ident}[1]{\mathbb{I}_{#1}}
\newcommand{\iden}{\mathbb{I}}
\newcommand{\flip}[1]{{#1}^\smallsmile}
\newcommand{\kleeneplus}[1]{{#1}^+}
\newcommand{\kleenestar}[1]{{#1}^*}
\newcommand{\cmpl}[1]{\overline{#1}}
\newcommand{\rel}{\times}
\newcommand{\compose}{;}
\newcommand{\subs}{\dashv}
\newcommand{\fun}{\rightarrow}
\newcommand{\N}{\mbox{\msb N}}
\newcommand{\disjn}[1]{\id{disjoint}(#1)}
\newcommand{\fsignat}[3]{\id{#1}:\id{#2}\mbox{$\rightarrow$}\id{#3}}
\newcommand{\signat}[3]{\mbox{${#1}_{[{#2},{#3}]}$}}
\newcommand{\declare}[3]{\id{#1}:\id{#2}\mbox{$\times$}\id{#3}}

\title{AMMBR: A Method to Maintain Business Rules}
\author{Stef Joosten}
\begin{document}
\maketitle
\begin{abstract}
	Ideally, an information infrastructure supports an organization by maintaining rules that are specific to the business.
	Such rules are known as \define{business rules} \cite{BRM}.
	This paper proposes a method to define such an infrastructure.
	It allows organizations to build information systems that comply to their business rules in a provable way.
	The method is implemented in a tool that generates functional specifications.
	The tool is an instrument for designers (information architects) who define processes and information systems.

	This paper is meant for readers who are interested in the mathematical background of
	deriving software from business rules.
	It uses a relation algebra as a vehicle for processing business rules.
\end{abstract}
keywords: compliance, business rule, relation algebra, software generation, provable software, process control.
\section{Introduction}
	In order to define an information system to support an organization,
	we look at organizations as systems that maintain rules.
	An ideal information infrastructure supports employees and other stakeholders to maintain the rules of the business.
	Yet rules must be concrete if they are to be used by computers.
	The method is therefore restricted to those rules of which
	violation can be established unambiguously and objectively.
	To {\em maintain} a rule means to prevent or correct all violations, either by human or automated actions.
	Temporary violation of a rule is only acceptable for rules that are maintained by humans,
	but the infrastructure must signal these violations and make them available to those people

@@

\section{Clauses}
\label{sct:Clauses}
	This section explains:
\begin{enumerate}
\item	how to derive clauses from a set of rules, by a normalization procedure.
\item	how to transform clauses into left identity clauses and right identity clauses, if possible.
\item	how to derive multiplicity properties from clauses thus obtained.
\end{enumerate}
	A \define{rule} is an expression $e$ that must be maintained.
	To \define{maintain} rule $r$ means to ensure that $\full\subs r$ at all times,
	where $\full$ is the complete relation (i.e. everything is an element of $\full$).
	In order to discuss maintaining rules, we must introduce some vocabulary.
	As long as $\full\subs r$, we say that $r$ is \define{satisfied}, or simply that it \define{holds}.
	If $r$ is not $\full$, we say that \define{invariance is broken},
	or rule $r$ is \define{not satisfied}, or it \define{does not hold}.
	Since $r\cup\cmpl{r}=\full$, the missing pairs are determined by $\cmpl{r}$.
	Each element of $\cmpl{r}$ is called a \define{violation} of $r$.
	To \define{restore invariance} of rule $r$ means to change the contents of $r$ in such a way that $\full\subs r$.

@@

\section{Code Fragments}
	In this section we derive code for computers,
	which is built around two basic operations: insert and delete.
	The insert operation can be represented in terms of pre- and postconditions\footnote{this idea
	was introduced by Floyd and Hoare.
	It is well known for making assertions about computer programs and reasoning about them.}.
\[\begin{array}{r@{~}l}
&\{\text{pre: $\code{P}=p$}\}\\
\code{INSERT}&\Delta\code{ IN P}\\
&\{\text{post: $\code{P}=p\cup\Delta$}\}\\
\end{array}\]
	The same can be expressed mathematically by the function $\id{insert}_\code{P}$:
\begin{equation}
\id{insert}_\code{P}(\Delta,r)\ =\ \subst{P}{P\cup\Delta}{r}
\end{equation}
	This function inserts a set of pairs, $\Delta$, into its argument.

	The delete operation is introduced similarly:
\[\begin{array}{r@{~}l}
&\{\text{pre: $\code{P}=p$}\}\\
\code{DELETE}&\Delta\code{ FROM P}\\
&\{\text{post: $\code{P}=p\cap\cmpl{\Delta}$}\}\\
\end{array}\]
	This too is expressed mathematically by the function $\id{delete}_\code{P}$:
\begin{equation}
\id{delete}_\code{P}(\Delta,r)\ =\ \subst{P}{P\cap\cmpl{\Delta}}{r}
\end{equation}

	Let one example introduce the idea of breaking and restoring invariance.
	Suppose the rule $\code{P}\subs\code{S}\cup\code{T}$ is to be maintained by a computer,
	and \code{P}, \code{S}, and \code{T} are variables (memory locations) in that computer,
	which represent relations.
	We assume that before executing any code, we have $\code{P}=p$, $\code{S}=s$, and $\code{T}=t$.
	The following example shows that inserting a set of pairs, $\Delta$, in \code{P}
	may violate the rule.
	Inserting the same $\Delta$ in \code{S} restores invariance.
	In terms of code, we write:
\[\begin{array}{r@{~}l}
&\{\text{Assume $\code{P}=p$, $\code{S}=s$, $\code{T}=t$ and $\code{P}\subs\code{S}\cup\code{T}$}\}\\
\code{INSERT}&\Delta\code{ IN P}\\
&\{\text{$\code{P}=p\cup\Delta$, $\code{S}=s$, and $\code{T}=t$, and invariance is broken: $\code{P}\not\subs\code{S}\cup\code{T}$}\}\\
\code{INSERT}&\Delta\code{ IN S}\\
&\{\text{postcondition: $\code{P}=p\cup\Delta$, $\code{S}=s\cup\Delta$, $\code{T}=t$, and $\code{P}\subs\code{S}\cup\code{T}$}\}\\
\end{array}\]
	To discuss this example, we start at the top.
	We may assume $\code{P}\subs\code{S}\cup\code{T}$ that holds.
	Rephrased in disjunctive form, this is equivalent to $\cmpl{\code{P}}\cup\code{S}\cup\code{T}$.
	The first line of code inserts a set of pairs, $\Delta$, into relation \code{P},
	which means that after the first line, we have $\code{P}=p\cup\Delta$.
	Consequently, $\cmpl{\code{P}}=\cmpl{p}\cap\cmpl{\Delta}$,
	which means that some elements of $\Delta$ may be missing from the rule, $\cmpl{\code{P}}\cup\code{S}\cup\code{T}$.
	To neutralize that result,
	the missing elements must be added to either S or T.
	That is what happens in the second line.

	To prove that rule $\code{P}\subs\code{S}\cup\code{T}$ is maintained,
	we must show that its value afterwards is equal to $\full$,
	under the condition that its initial value equals $\full$.
	The initial value is $\cmpl{p}\cup s\cup t$, which we may assume.
	The final value is $(p\cup\Delta)\subs(s\cup\Delta)\cup t$, which must be proven equal to $\full$:
\[\begin{array}{cl}
&(p\cup\Delta)\subs(s\cup\Delta)\cup t\\
=&\hspace{1in}\{\hbox{by law \ref{eqn:rid of subseteq}}\}\\
&\cmpl{p\cup\Delta}\ \cup\ s\ \cup\ \Delta\ \cup\ t\\
=&\hspace{1in}\{\hbox{De Morgan's law (\ref{eqn:De Morgan})}\}\\
&(\cmpl{p}\cap\cmpl{\Delta})\ \cup\ s\ \cup\ \Delta\ \cup\ t\\
=&\hspace{1in}\{\hbox{Distribute $\cup$ over $\cap$}\}\\
&(\cmpl{p}\cup s\cup\Delta\cup t)\ \cap\ (\cmpl{\Delta}\cup s\cup\Delta\cup t)\\
=&\hspace{1in}\{\cmpl{\Delta}\cup\Delta=\full\}\\
&\cmpl{p}\cup s\cup\Delta\cup t\\
=&\hspace{1in}\{\full\ =\ \cmpl{p}\cup s\cup t\}\\
&\full
\end{array}\]
	From this derivation we may conclude that invariance of $\code{P}\subs\code{S}\cup\code{T}$ is restored.

@@

	Each of the code fragments from table \ref{tbl:code fragments} will maintain invariance of $e$.
	We shall prove this for one code fragment only (we have arbitrarily picked the third code fragment):
	This gives
\[\begin{array}{rl}
&\cmpl{(p\cup\Delta)}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ (t\cup\cmpl{a(e)})\\
=&~~~~~~~~\{a(e)\ =\ \cmpl{(p\cup\Delta)}\cup\cmpl{q}\cup s\cup t\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{\cmpl{(p\cup\Delta)}\cup\cmpl{q}\cup s\cup t}\\
=&~~~~~~~~\{\text{De Morgan (\ref{eqn:De Morgan})}\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ (\cmpl{\cmpl{(p\cup\Delta)}}\cap\cmpl{\cmpl{q}}\cap\cmpl{s}\cap\cmpl{t})\\
=&~~~~~~~~\{\text{remove double complement}\}\\
&\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ ((p\cup\Delta)\cap q\cap\cmpl{s}\cap\cmpl{t})\\
=&~~~~~~~~\{\text{Distribute $\cap$ over $\cup$}\}\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ (p\cup\Delta))\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ q)\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{s})\ \cap\\
&(\cmpl{p\cup\Delta}\ \cup\ \cmpl{q}\ \cup\ s\ \cup\ t\ \cup\ \cmpl{t})\\
=&~~~~~~~~\{\cmpl{p\cup\Delta}\ \cup\ p\cup\Delta=\full\text{~and~}\cmpl{q}\cup q=\full\text{~and~}s\ \cup\ \cmpl{s}=\full\text{~and~}t\ \cup\ \cmpl{t}=\full\}\\
&\full\cap \full\cap \full\cap \full\\
=\\
&\full\\
\end{array}\]
